### 2022-08-08 chapters/optimization/python-dir/python.tex

\subsection{Python and machine learning }
Machine learning is becoming an integral part of our daily lives, growing more potent and energy-hungry each year.

@@

\paragraph{hardware }
Chifflot 8 from Grid 5000's Lille site was used for all of the trials.
The machine is outfitted with two Intel Xeon Gold 6126 CPUs, each having 12 physical cores, 192 GB of RAM, and two 32 GB Tesla V100 GPUs.

\paragraph{software}

For the sake of reproducibility, each experiment is done within a Docker container using Jupyter lab . These tests are run on top of a minimal version of Debian-10 in order to increase the accuracy of the tests by elimanting any unnecessary processes.

\subsubsection{Workload}
\paragraph{Models}

Several models were developed, however only two were used in the final trials because they achieved 94 percent accuracy in an acceptable length of time.
David Page's cifar10-fast and Woonhyuk Baek's torchskeleton are shown here.
\paragraph{Datasets}

The CIFAR-10 dataset was the major source of data for the studies.
It is made up of 60000 32x32 color images grouped into ten categories.

@@

\subsubsection{candidates}
The experiments were run with several different CPU and GPU configurations:

\begin{itemize}
    \item With and without GPU
    \item With and without hyperthreading
    \item Different number of physical cores
\end{itemize}

\subsubsection{metrics}

We used Pytorch 1.10.0  to train those models and pyjoules to measure the energy consumption of the GPU and CPU.
\begin{itemize}
    \item accuracy : in \%
    \item execution time : in seconds for both the duration of each epoch and the total duration to a chieve a certain accuracy
    \item total energy consumption : for the CPU and the  GPU
    \item
\end{itemize}

\section{finding and results}
As the model's accuracy increases, so does the energy required for the next accuracy increment.

@@

\paragraph{conclusion}
We discovered that when a model's accuracy improves, so does the energy required for a subsequent accuracy gain.
This begs the question of when we should discontinue training.
Is a 10\% increase in accuracy worth it if we have to spend three times the energy?
Even while utilizing a GPU consumes more total power, the training time is significantly reduced.
As a result, the utilization of a GPU is required to lower the energy consumption of the training.
Some experiments revealed unusual behavior that was not studied during this internship.
These could be examined in future studies and perhaps used to achieve more energy savings.
The execution time did not depend on the number of cores employed in several studies.
This parameter had a significant impact on training time in others.
As a result, the best core configuration will be determined by the script.
The next steps in this area could include exploring the reasons for these discrepancies and possibly creating a script that can automatically find the ideal core configuration for a specific script.

>>>>

\subsection{Python \& Machine Learning}
Machine learning is becoming an integral part of our daily lives, growing more potent and energy-hungry each year.

@@

\paragraph{Hardware settings}
Chifflot 8 from Grid 5000's Lille site was used for all of the trials.
The machine is outfitted with two Intel Xeon Gold 6126 CPUs, each having 12 physical cores, 192 GB of RAM, and two 32 GB Tesla V100 GPUs.

\paragraph{Software settings}
For the sake of reproducibility, each experiment is done within a Docker container using Jupyter lab. These tests are run on top of a minimal version of Debian-10 to increase the accuracy of the tests by eliminating any unnecessary processes.

\subsubsection{Input Workload}
\paragraph{Models}
Several models were developed, however, only two were used in the final trials because they achieved 94 percent accuracy in an acceptable length of time.
David Page's cifar10-fast and Woonhyuk Baek's torch skeleton are shown here.

\paragraph{Datasets}
The CIFAR-10 dataset was the major source of data for the studies.
It is made up of 60000 32x32 color images grouped into ten categories.

@@

\subsubsection{candidates}
The experiments were run with several different CPU and GPU configurations:
\begin{itemize}
    \item with and without GPU,
    \item with and without CPU hyper-threading,
    \item different number of CPU physical cores.
\end{itemize}

\subsubsection{Key Performance Metrics}
We used Pytorch 1.10.0 to train those models and pyjoules to measure the energy consumption of the GPU and CPU.
\begin{itemize}
    \item accuracy : in \%
    \item execution time : in seconds for both the duration of each epoch and the total duration to a chieve a certain accuracy
    \item total energy consumption : for the CPU and the  GPU
    \item
\end{itemize}

\section{Results \& Findings}
As the model's accuracy increases, so does the energy required for the next accuracy increment.

@@

\paragraph{Conclusion}
We discovered that when a model's accuracy improves, so does the energy required for a subsequent accuracy gain.
This begs the question of when we should discontinue training.
Is a 10\% increase in accuracy worth it if we have to spend three times the energy?
Even while utilizing a GPU consumes more total power, the training time is significantly reduced.
As a result, the utilization of a GPU is required to lower the energy consumption of the training.
Some experiments revealed unusual behavior that was not studied during this internship.
These could be examined in future studies and perhaps used to achieve more energy savings.
The execution time did not depend on the number of cores employed in several studies.
This parameter had a significant impact on training time in others.
As a result, the best core configuration will be determined by the script.
The next steps in this area could include exploring the reasons for these discrepancies and possibly creating a script that can automatically find the ideal core configuration for a specific script.

### 2022-08-08 chapters/programming_languages/rpc/rpc.tex

\subsection*{introduction}
With the emergence of cloud technologies, many protocols wanted to take the lead.
Nowdays, most of the architectures are based on multi-services and microservices.
And, to have higher versatility of developers multiple companies choose to be open to different programming languages.
Basically, it would be more efficient if we take advantage of each programming language to satisfy a specific need.
However, the challenge nowadays is to make the bridge between those platforms.
We have many initiatives, such as OpenAPI, that try to create a taxonomy for RESTful APIs.
Other approaches implement all the different interfaces of the protocol by themselves, such as RPC.

@@

\subsection{Experimental protocol}
\subsubsection{measurement context}
\paragraph{Hardware settings}
All the experiments are run on the cluster \textsf{paravance} of the G5K platform.
This cluster is composed of 72 identical machines, each one is equipped with 2 Intel Xeon E5-2630 V3, with 128\,GB of RAM.
For more accuracy, our SUT (\emph{System Under Test}) is equipped with a minimal version of Debian\,9 (4.9.0 kernel version), which enforces the core processes required for the purpose of our experiment.
Furthermore, we used Docker containers technology for reproducibility of the experiments and the isolation of the servers.
\paragraph{Client and server environments}
To limit the impact of the network on the experiments, we run both the client and the server on the same machine.
However, we isolate each part on a separate socket, in order to reduce the effect that the client might have on the server and \emph{vise versa}.
To do so, for each iteration, we always run the same client on \textsf{socket\,0} and the server that we want to test on \textsf{socket\,1}.
Both the server and the client use the whole socket for their experiment.
In addition, all the additional services, such as the kernel and HwPC sensor, are run on \textsf{socket\,0}.
Therefore, the only process being executed in \textsf{socket\,1} is the server that we benchmark and monitor.

\subsubsection{metrics}
\paragraph{Energy measurements}
To report on the energy consumption, we used HwPC sensor [TOCITE], which is based on Intel RAPL technology, one of the most accurate tools to measure the energy consumption of the CPU and DRAM [TOCITE].
For better accuracy, we ran the HwPC sensor with a frequency of 10~Hz, and we used the same machine for all the experiments in order to reduce the variability [TOCITE].

@@

\subsubsection{workload}
The purpose of the experiment is to analyse the behaviour of different GRPC implementations. therefore we have two kind of worlkoads 
\paragraph{number of clients}: this is the number of concurrent clients that we want to test. It is automatically handled with the ghz client 
\paragraph{payload}: the size of each request . this varries from 50 Bytes up to 10 MB.

@@

\subsection{Results and findings}
[\textsc{RQ}~1:] How do RPC implementations react to the size of the data ?
The purpose of this question is to study the behaviour of the server when transferring large objects.
To do so, we send 80,000 requests to the server whose size scales from 10 bytes up to 10 Megabytes, which results in 10,000 requests per size per server.
To eliminate extra factors, we let the server handle the rate at which it can answer each request.
However, we put a 20 seconds timeout limit for each request.
Therefore, our boundary condition is only the number of requests received by the server.
For this experiment, we investigate 4 observable variables~:
\begin{enumerate}
    \item the \textsf{average power consumption} during the the process: this will indicate the overall behaviour of the server in working mode for long durations,
    \item the \textsf{tail latency} for the 99th percentile, which indicates how performant is the server,
    \item the \textsf{average number of requests per second}, which indicates the average number of clients that the server can handle,
    \item the \textsf{average energy cost} of a single request: unlike the first indicator, this one shows how green is the implementation taking performance into consideration.
\end{enumerate}

@@

\subsubsection{Stress free mode}
In this mode, the compiled languages tend to consume less resources (average power).
JVM-based languages tend to consume more energy, especially Scala.
However, we do not observe the same behaviour when it comes to efficiency.
Unlike the other interpreted programming languages, PHP performances could be compared to the compiled ones, such as CPP or GO, and even better to some others, such as Swift.
JVM-based languages tend to have better performances than the interpreted ones.
Furthermore, OpenJDK has shown more efficiency than GraalVM [].
Overall, we can have 3 groups when it comes the cost of each request:
\begin{itemize}
    \item Energy-efficient class: C++, GO, RUST, ELIXIR, and PHP,
    \item Middle class: Most of the interpreted languages and VM-based ones,
    \item Energy-greedy class: Crystal and Scala.
\end{itemize}

\subsubsection{Escalation mode}
In this mode, the behaviour of the server depends on the payload. We observe three behaviours:
\begin{enumerate}
    \item Drop in performances without an increased power, such as .Net core, Java micronaut, Crystal, and Dart.
          In this case, the server keeps using the same resources, and sometimes less, because it takes more time to handle the less requests.
          This class of languages tend to be the most energy consuming when it comes to the cost per request,
    \item Increase in power without affecting the performances: such as Go, .Net.
          The energy consumption of a single request, is affected slightly but still increase,
    \item Increase in power and drop in performances: Despite the increase of the power consumption, the server becomes slightly slower, which increases the cost of the energy cost per request.
          This cost is still better than the first case, which concludes that the servers in the first category are on the verge of breaking.
\end{enumerate}

@@

\subsection{Threads to validity}

>>>>

\subsection{Introduction}
With the emergence of cloud technologies, many protocols wanted to take the lead.
Nowdays, most of the architectures are based on multi-services and microservices.
And, to have higher versatility of developers multiple companies choose to be open to different programming languages.
Basically, it would be more efficient if we take advantage of each programming language to satisfy a specific need.
However, the challenge nowadays is to make the bridge between those platforms.
We have many initiatives, such as OpenAPI, that try to create a taxonomy for RESTful APIs.
Other approaches implement all the different interfaces of the protocol by themselves, such as RPC.

@@

\subsection{Experimental Protocol}
\subsubsection{Measurement Context}
\paragraph{Hardware settings}
All the experiments are run on the cluster \textsf{paravance} of the G5K platform.
This cluster is composed of 72 identical machines, each one is equipped with 2 Intel Xeon E5-2630 V3, with 128\,GB of RAM.
For more accuracy, our SUT (\emph{System Under Test}) is equipped with a minimal version of Debian\,9 (4.9.0 kernel version), which enforces the core processes required for the purpose of our experiment.
Furthermore, we used Docker containers technology for reproducibility of the experiments and the isolation of the servers.
\paragraph{Client and server environments}
To limit the impact of the network on the experiments, we run both the client and the server on the same machine.
However, we isolate each part on a separate socket, in order to reduce the effect that the client might have on the server and \emph{vise versa}.
To do so, for each iteration, we always run the same client on \textsf{socket\,0} and the server that we want to test on \textsf{socket\,1}.
Both the server and the client use the whole socket for their experiment.
In addition, all the additional services, such as the kernel and HwPC sensor, are run on \textsf{socket\,0}.
Therefore, the only process being executed in \textsf{socket\,1} is the server that we benchmark and monitor.

\subsubsection{Key Performance Metrics}
\paragraph{Energy measurements}
To report on the energy consumption, we used HwPC sensor [TOCITE], which is based on Intel RAPL technology, one of the most accurate tools to measure the energy consumption of the CPU and DRAM [TOCITE].
For better accuracy, we ran the HwPC sensor with a frequency of 10~Hz, and we used the same machine for all the experiments in order to reduce the variability [TOCITE].

@@

\subsubsection{Workload}
The purpose of the experiment is to analyse the behaviour of different GRPC implementations. therefore we have two kind of worlkoads 
\paragraph{number of clients}: this is the number of concurrent clients that we want to test. It is automatically handled with the ghz client 
\paragraph{payload}: the size of each request . this varries from 50 Bytes up to 10 MB.

@@

\subsection{Results \& Findings}
[\textsc{RQ}~1:] How do RPC implementations react to the size of the data ?
The purpose of this question is to study the behaviour of the server when transferring large objects.
To do so, we send 80,000 requests to the server whose size scales from 10 bytes up to 10 Megabytes, which results in 10,000 requests per size per server.
To eliminate extra factors, we let the server handle the rate at which it can answer each request.
However, we put a 20 seconds timeout limit for each request.
Therefore, our boundary condition is only the number of requests received by the server.
For this experiment, we investigate 4 observable variables~:
\begin{enumerate}
    \item the \textsf{average power consumption} during the the process: this will indicate the overall behaviour of the server in working mode for long durations,
    \item the \textsf{tail latency} for the 99th percentile, which indicates how performant is the server,
    \item the \textsf{average number of requests per second}, which indicates the average number of clients that the server can handle,
    \item the \textsf{average energy cost} of a single request: unlike the first indicator, this one shows how green is the implementation taking performance into consideration.
\end{enumerate}

@@

\subsubsection{Stress Free Mode}
In this mode, the compiled languages tend to consume less resources (average power).
JVM-based languages tend to consume more energy, especially Scala.
However, we do not observe the same behaviour when it comes to efficiency.
Unlike the other interpreted programming languages, PHP performances could be compared to the compiled ones, such as CPP or GO, and even better to some others, such as Swift.
JVM-based languages tend to have better performances than the interpreted ones.
Furthermore, OpenJDK has shown more efficiency than GraalVM [].
Overall, we can have 3 groups when it comes the cost of each request:
\begin{itemize}
    \item Energy-efficient class: C++, GO, RUST, ELIXIR, and PHP,
    \item Middle class: Most of the interpreted languages and VM-based ones,
    \item Energy-greedy class: Crystal and Scala.
\end{itemize}

\subsubsection{Escalation Mode}
In this mode, the behaviour of the server depends on the payload. We observe three behaviours:
\begin{enumerate}
    \item Drop in performances without an increased power, such as .Net core, Java micronaut, Crystal, and Dart.
          In this case, the server keeps using the same resources, and sometimes less, because it takes more time to handle the less requests.
          This class of languages tend to be the most energy consuming when it comes to the cost per request,
    \item Increase in power without affecting the performances: such as Go, .Net.
          The energy consumption of a single request, is affected slightly but still increase,
    \item Increase in power and drop in performances: Despite the increase of the power consumption, the server becomes slightly slower, which increases the cost of the energy cost per request.
          This cost is still better than the first case, which concludes that the servers in the first category are on the verge of breaking.
\end{enumerate}

@@

\subsection{Threads to Validity}
